---
title: "Other models"
author: "Yesol Sapozhnikov"
date: "5/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose of this script

Apply the same steps to determine models for different combination of datasets (FoldX, or MD + FoldX) and sets of parameters (FoldX energy terms, biochemical/physical properties, MD-associated SD).


## Read in data, load packages

```{r}
library(tidyverse)
library(leaps)

source("functions.R")
```

```{r}
# Full tables:
fold <- read.csv("../outputs/table_full_f.csv")
bind <- read.csv("../outputs/table_full_b.csv")
fold_tr <- read.csv("../outputs/tr_table_full_f.csv")
bind_tr <- read.csv("../outputs/tr_table_full_b.csv")

# FoldX-only tables:
foldx <- read.csv("../outputs/table_fxonly_f.csv")
bindx <- read.csv("../outputs/table_fxonly_b.csv")
foldx_tr <- read.csv("../outputs/tr_table_fxonly_f.csv")
bindx_tr <- read.csv("../outputs/tr_table_fxonly_b.csv")
```

There were 3 rows in folding dataset that had no match for dssp data and removed from "fold" and "fold_tr." Remove these manually from "foldx" and "foldx_tr." (Run 01_generate-organized-tables.Rmd to get details on the missing rows.)
```{r}
foldx <- foldx[-c(449, 642, 643),]
foldx_tr <- foldx_tr[-c(449, 642, 643),]

# Make sure everything else matches
sum(fold$mut == foldx$mut)
```




## Intervening model 1: MD+FX dataset with FX energy terms only

Get models from stepwise selection and best subset for each dataset.
```{r, include = FALSE}
data_f <- fold_tr %>% select(backbone_vdwclash:error)
data_b <- bind_tr %>% select(backbone_vdwclash:error)
# stepwise selection
mod1_step_f <- run_stepwise(data_f)
mod1_step_b <- run_stepwise(data_b)
# best subset
mod1_subset_f <- run_subset(data_f)
mod1_subset_b <- run_subset(data_b)
```

Cross validation.
```{r}
models <- list(fold_stepwise = mod1_step_f, fold_best_subset = mod1_subset_f,
               bind_stepwise = mod1_step_b, bind_best_subset = mod1_subset_b)
summary_tab <- data.frame(models = names(models))
datasets_ls <- list(fold, fold, bind, bind)
for (i in 1:length(models)){
  summary_tab$adj_r2[i] <- summary(models[[i]])$adj.r.squared
  out <- loocv(formula(models[[i]]), datasets_ls[[i]])
  summary_tab$coverage[i] <- out$coverage
  summary_tab$med_upr_bd[i] <- out$med_upr_bd
}
summary_tab

```
For folding, best subset model is better. For binding, both methods produced same models.
```{r}
mod1_subset_f
mod1_subset_b
```


## MD+FX dataset with FX energy terms, biochemical predictors

Get models from stepwise selection and best subset for each dataset.
```{r, include = FALSE}
data_f <- fold_tr %>% select(-ends_with("sd"))
data_b <- bind_tr %>% select(-ends_with("sd"))
# stepwise selection
mod2_step_f <- run_stepwise(data_f)
mod2_step_b <- run_stepwise(data_b)
# best subset
mod2_subset_f <- run_subset(data_f)
mod2_subset_b <- run_subset(data_b)
```

Cross validation.
```{r}
models <- list(fold_stepwise = mod2_step_f, fold_best_subset = mod2_subset_f,
               bind_stepwise = mod2_step_b, bind_best_subset = mod2_subset_b)
summary_tab <- data.frame(models = names(models))
datasets_ls <- list(fold, fold, bind, bind)
for (i in 1:length(models)){
  summary_tab$adj_r2[i] <- summary(models[[i]])$adj.r.squared
  out <- loocv(formula(models[[i]]), datasets_ls[[i]])
  summary_tab$coverage[i] <- out$coverage
  summary_tab$med_upr_bd[i] <- out$med_upr_bd
}
summary_tab
```
Best subset models are better in both cases.

```{r}
mod2_subset_f
mod2_subset_b
```



## FX dataset with energy terms and biochemical predictors

Get models from stepwise selection and best subset for each dataset.
```{r, include = FALSE}
data_f <- foldx_tr %>% select(-ends_with("sd"))
data_b <- bindx_tr %>% select(-ends_with("sd"))
# stepwise selection
mod3_step_f <- run_stepwise(data_f)
mod3_step_b <- run_stepwise(data_b)
# best subset
mod3_subset_f <- run_subset(data_f)
mod3_subset_b <- run_subset(data_b)
```

Cross validation.
```{r}
models <- list(fold_stepwise = mod3_step_f, fold_best_subset = mod3_subset_f,
               bind_stepwise = mod3_step_b, bind_best_subset = mod3_subset_b)
summary_tab <- data.frame(models = names(models))
datasets_ls <- list(foldx, foldx, bindx, bindx)
for (i in 1:length(models)){
  summary_tab$adj_r2[i] <- summary(models[[i]])$adj.r.squared
  out <- loocv(formula(models[[i]]), datasets_ls[[i]])
  summary_tab$coverage[i] <- out$coverage
  summary_tab$med_upr_bd[i] <- out$med_upr_bd
}
summary_tab
```
For folding dataset, best subset model is slightly better. Both methods produce same model for binding. 

```{r}
mod3_subset_f
mod3_subset_b
```






