---
title: "Figures"
author: "Yesol Sapozhnikov"
date: '2022-06-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get data frames for plotting

Load libraries and build data frames for plots.
```{r}
library(tidyverse)
library(ggpubr)
library(ggpattern)
#library(ggExtra)

# Full tables:
fold <- read.csv("../outputs/table_full_f.csv")
bind <- read.csv("../outputs/table_full_b.csv")
# FoldX-only tables:
foldx <- read.csv("../outputs/table_fxonly_f.csv")
bindx <- read.csv("../outputs/table_fxonly_b.csv")
# selected models from stepwise & best subset selection
mods <- readRDS("../outputs/models.rds")
# source LOsysOCV function
source("../scripts/functions.R")

# Run LOOCV to create dataframes
loo_fold <- loocv(formula(mods$mod_subset_f), fold)$df
loo_bind <- loocv(formula(mods$mod_subset_b), bind)$df
```

## Table: System-by-system breakdown

```{r}
loo_fold %>% group_by(systems) %>%
  summarise(n = n(),
            median_error = median(error),
            coverage = round(mean(test),2),
            width = median(upr))
```

```{r}
loo_bind %>% group_by(systems) %>%
  summarise(n = n(),
            median_error = median(error),
            coverage = round(mean(test),2),
            width = median(upr))
```


## Prediction interval plot with "misses"

Overall plots
```{r}
errors_all_fold <- loo_fold %>% arrange(upr) %>%
  ggplot(aes(x = seq_along(error), y = error)) +
  geom_point(aes(color = test)) +
  geom_line(aes(x = seq_along(error), y = upr, color = "upper bound")) +
  scale_color_manual(values = c("red", "black", "blue"),
                     labels = c("error above bound", "error within bound","predicted upper bound"),
                     guide = guide_legend(override.aes = list(
                       linetype = c("blank","blank","solid"),
                       shape = c(16,16,NA)
                     ))) +  
  theme_light() +
  scale_x_continuous(NULL, breaks = NULL) +
  labs(y = "error (kcal/mol)") +
  annotate("text", x = 100, y = 12.5, label = "Fold data", size = 8) +
  theme(legend.position = c(0.7,0.8), legend.title = element_blank())
errors_all_bind <- loo_bind %>% arrange(upr) %>%
  ggplot(aes(x = seq_along(error), y = error)) +
  geom_point(aes(color = test)) +
  geom_line(aes(x = seq_along(error), y = upr, color = "upper bound")) +
  scale_color_manual(values = c("red", "black", "blue"),
                     labels = c("error above bound", "error within bound","predicted upper bound"),
                     guide = guide_legend(override.aes = list(
                       linetype = c("blank","blank","solid"),
                       shape = c(16,16,NA)
                     ))) +  
  theme_light() +
  scale_x_continuous(NULL, breaks = NULL) +
  labs(y = "error (kcal/mol)") +
  annotate("text", x = 90, y = 15.7, label = "Bind data", size = 8) +
  theme(legend.position = c(0.7,0.8), legend.title = element_blank())
```

Breakdown by system
```{r}
errors_sys_fold <- loo_fold %>% arrange(upr) %>%
  ggplot(aes(y=error,x=seq_along(error))) +
  geom_point(aes(color=test), size = .8, show.legend = FALSE) +
  geom_line(aes(x=seq_along(error), y=upr), color="blue") +
  scale_color_manual(values = c("#FF0000", "#000000")) +
  facet_wrap(~systems, nrow = 2) +
  theme_bw() + labs(y = "error (kcal/mol)") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"))
errors_sys_bind <- loo_bind %>% arrange(upr) %>% 
  ggplot(aes(y=error,x=seq_along(error))) +
  geom_point(aes(color=test), size = .8, show.legend = FALSE) +
  geom_line(aes(x=seq_along(error), y=upr), color="blue") +
  scale_color_manual(values = c("#FF0000", "#000000")) +
  theme_bw() + labs(y = "error (kcal/mol)") +
  facet_wrap(~systems, nrow = 2) +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"))
```

Arragne plots
```{r}
ggarrange(errors_all_fold, errors_sys_fold, errors_all_bind, errors_sys_bind,
          labels = c("A","B","C","D"),
          ncol = 2, nrow = 2)
ggsave("../figures/errors_sorted.png", width = 11, height = 9)
```


## Correlation between ddG and error

```{r}
ggplot(loo_fold, aes(x = ddg_exp, y = error)) +
  geom_point(aes(color = test)) +
  labs(x = "ddG", y = "error")
```
```{r}
ggplot(loo_bind, aes(x = ddg_exp, y = error)) +
  geom_point(aes(color = test)) +
  labs(x = "ddG", y = "error")
```




## Predicted stability effects vs actual effects

If we classify stability effects according to ddG, how concordant are they between actual and predicted?
Cut off at +/- 0.5 kcal/mol as in Khan & Vihinen.

```{r}
binned <- loo_fold %>% 
  select(systems, mut, ddg_exp, total, error, test) %>%
  rename(ddg_pred = total) %>%
  mutate(classes_pred = cut(ddg_pred, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")),
         classes_actual = cut(ddg_exp, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")))
ggplot(binned, aes(fill = classes_actual, x = classes_pred)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("blue", "darkgray", "red")) +
  theme_classic()
```

```{r}
binned_bind <- loo_bind %>% 
  select(systems, mut, ddg_exp, total, error, test) %>%
  rename(ddg_pred = total) %>%
  mutate(classes_pred = cut(ddg_pred, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")),
         classes_actual = cut(ddg_exp, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")))
ggplot(binned_bind, aes(fill = classes_actual, x = classes_pred)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("blue", "darkgray", "red")) +
  theme_classic()
```

Combine fold and bind altogether
```{r}
binned_all <- bind_rows(binned, binned_bind)
# Re-order labels so that figure is more intuitive
# binned_all$classes_pred <- factor(binned_all$classes_pred,
#                                   levels = c("destabilizing", "neutral", "stabilizing"))
binned_all$classes_actual <- factor(binned_all$classes_actual,
                                  levels = c("destabilizing", "neutral", "stabilizing"))

classification_error <- 
ggplot(binned_all, aes(fill = classes_actual, x = classes_pred)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("red", "darkgray", "blue"), name = "Actual class") +
  theme_classic() +
  labs(x = "Predicted class", y = "Frequency")
classification_error
#ggsave("../figures/classification_error.png", width = 7, height = 5)
```


Scatter plot
```{r}
error_correlation <-
binned_all %>%
  ggplot(aes(x = ddg_pred, y = ddg_exp)) +
  geom_point(aes(color = error)) +
  theme_classic() +
  scale_color_gradient(low = "blue", 
                        high = "red") +
  geom_point(data = binned_all[binned_all$test == FALSE,], 
             aes(x = ddg_pred, y = ddg_exp, color = error),
             shape = 4, size = 4) +
  labs(x = expression(paste(Delta, Delta, "G")[pred]),
       y = expression(paste(Delta, Delta, "G")[exp]))
error_correlation
```

```{r}
ggarrange(error_correlation, classification_error,
          labels = c("A","B"),
          ncol = 2)
ggsave("../figures/predicted_vs_actual.png", width = 12, height = 5)
```
```{r}
error_correlation <- 
binned_all %>%
  ggplot(aes(x = ddg_exp, y = ddg_pred)) +
  geom_point(aes(color = error)) +
  theme_bw() +
  scale_color_gradient(low = "chartreuse", 
                        high = "red4") +
  geom_point(data = binned_all[binned_all$test == FALSE,], 
             aes(x = ddg_exp, y = ddg_pred, color = error),
             shape = 4, size = 4) +
  labs(x = expression(paste(Delta, Delta, "G")[exp]~(kcal/mol)),
       y = expression(paste(Delta, Delta, "G")[pred]~(kcal/mol))) +
  theme(legend.position = c(.85, .75))
```

Try histogram?

```{r}
classification <- 
binned_all %>% ggplot(aes(x = ddg_exp)) +
  geom_histogram(binwidth = 0.2, position = "dodge",
                 aes(fill = classes_pred)) +
  geom_vline(xintercept = c(-.5, .5), linetype = 2, size = .2) +
  scale_fill_manual(values = c("blue", "gray40", "red"), name = "Predicted class") +
  theme_bw() +
  theme(legend.position = c(.8, .6)) +
  labs(x = expression(paste(Delta, Delta, "G")[exp]~(kcal/mol)),
       y = "Number of mutations")
```




Density plot?
```{r}
classification_density <- 
ggplot(aes(x = ddg_exp, fill = classes_pred), data = binned_all) +
  geom_density(alpha = .5) +
  geom_vline(xintercept = c(-.5, .5), linetype = 2, size = .2) +
  scale_fill_manual(values = c("blue", "gray40", "red"), name = "Predicted class") +
  theme_bw() +
  theme(legend.position = c(.8, .6)) +
  labs(x = expression(paste(Delta, Delta, "G")[exp]~(kcal/mol)))
```


```{r}
ggarrange(error_correlation, classification_density,
          labels = c("A","B"),
          ncol = 2)
ggsave("../figures/predicted_vs_actual3.png", width = 12, height = 5)
```


More ideas about conveying the errors/misclassification.

```{r}
binned <- loo_fold %>% 
  select(systems, mut, ddg_exp, total, error, upr, test) %>%
  rename(ddg_pred = total) %>%
  mutate(classes_pred = cut(ddg_pred, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")),
         classes_actual = cut(ddg_exp, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")))
binned_bind <- loo_bind %>% 
  select(systems, mut, ddg_exp, total, error, upr, test) %>%
  rename(ddg_pred = total) %>%
  mutate(classes_pred = cut(ddg_pred, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")),
         classes_actual = cut(ddg_exp, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")))
binned_all <- bind_rows(binned, binned_bind)
# Re-order labels so that figure is more intuitive
# binned_all$classes_pred <- factor(binned_all$classes_pred,
#                                   levels = c("destabilizing", "neutral", "stabilizing"))
binned_all$classes_actual <- factor(binned_all$classes_actual,
                                  levels = c("destabilizing", "neutral", "stabilizing"))
```

```{r}
binned_all$ddg_pred_cat <- cut(binned_all$ddg_pred, breaks = c(-Inf, -1.5, -.5, .5, 1.5, 2.5, 4, 5.5, 10, Inf))
binned_all$bincode <- NA # to be filled with below loop
binned_all$class_match <- NA

for (i in 1: nrow(binned_all)){
  ddg <- binned_all$ddg_pred[i]
  L <- ddg - binned_all$upr[i]
  U <- ddg + binned_all$upr[i]
  if (L < -0.5 & U > 0.5){
    bincode <- "111"
  } else if (L > -0.5 & U < 0.5){
    bincode <- "010"
  } else if (U < -0.5){
    bincode <- "100"
  } else if (L > 0.5){
    bincode <- "001"
  } else if (U < 0.5){
    bincode <- "110"
  } else if (L > -0.5){
    bincode <- "011"
  }
  binned_all$bincode[i] <- bincode
  
  if (binned_all$classes_pred[i] == binned_all$classes_actual[i]){
    binned_all$class_match[i] <- "match"
  } else (binned_all$class_match[i] <- "mismatch")
}
```



```{r}
class1 <- 
binned_all %>% ggplot(aes(x = ddg_pred)) +
  geom_histogram_pattern(position = "fill", color = "black",
                 #breaks = c(-3, -2, -1, 0, 1, 2, 3, 4, 5, 6, 8, 10, 12, 16),
                 breaks = c(-2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 8, 10, 12, 16),
                 pattern_color = "black", pattern_density = 0.01, pattern_spacing = 0.02,
                 aes(fill = classes_actual, pattern = class_match)) +
  scale_pattern_manual(values = c("none", "stripe"), ) +
  scale_fill_manual(values = c("rosybrown1", "white", "lightsteelblue2"),
                    name = "Actual class",
                    labels = c("destabilizing", "neutral", "stabilizing")) +
  theme_classic() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.8, .6),
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[FoldX]~(kcal/mol)),
       y = "proportion") +
  guides(pattern = "none")
class1
```
```{r}
class2 <- 
binned_all %>% ggplot(aes(x = ddg_pred)) +
  geom_histogram(position = "fill", color = "black",
                 #breaks = c(-3, -2, -1, 0, 1, 2, 3, 4, 5, 6, 8, 10, 12, 16),
                 breaks = c(-2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 8, 10, 12, 16),
                 aes(fill = bincode)) +
  scale_fill_manual(values = c("red4", "rosybrown2", "lightsteelblue2", "white"),
                    name = "Possible class", 
                    labels = c("destabilizing", "destabilizing or neutral", "stabilizing or neutral", "any")) +
  theme_classic() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.75, .3), 
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[FoldX]~(kcal/mol)),
       y = "proportion")
class2
```

```{r}
ggarrange(class1, class2,
          labels = c("A","B"),
          ncol = 2)
ggsave("../figures/classifications.png", width = 11, height = 5)
```
```{r}
binned_all %>% ggplot(aes(x = ddg_pred)) +
  geom_histogram_pattern(position = "stack", color = "black", alpha = 0.2,
                 #breaks = c(-3, -2, -1, 0, 1, 2, 3, 4, 5, 6, 8, 10, 12, 16),
                 breaks = c(-2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 5),
                 pattern_color = "black", pattern_density = 0.01, pattern_spacing = 0.02,
                 aes(fill = classes_actual, pattern = class_match)) +
  scale_pattern_manual(values = c("none", "stripe"), ) +
  scale_fill_manual(values = c("red", "gray", "blue"),
                    name = "Actual class",
                    labels = c("destabilizing", "neutral", "stabilizing")) +
  theme_classic() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.8, .6),
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[FoldX]~(kcal/mol)),
       y = "proportion") +
  guides(pattern = "none")
```

```{r}
binned_all %>% ggplot(aes(x = ddg_exp)) +
  geom_histogram(position = "fill", color = "black",
                 #breaks = c(-3, -2, -1, 0, 1, 2, 3, 4, 5, 6, 8, 10, 12, 16),
                 breaks = c(-2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 8, 10, 12),
                 aes(fill = classes_pred)) +
  # scale_fill_manual(values = c("rosybrown1", "white", "lightsteelblue2"),
  #                   name = "Actual class",
  #                   labels = c("destabilizing", "neutral", "stabilizing")) +
  theme_classic() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.8, .6),
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[exp]~(kcal/mol)),
       y = "proportion") +
  guides(pattern = "none") + ggtitle("Predicted classes plotted along actual ddG")
```

## Resampling/simulation

```{r}
binned_fold <- loo_fold %>% 
  select(systems, mut, ddg_exp, total, error, upr, test) %>%
  rename(ddg_pred = total) %>%
  mutate(classes_pred = cut(ddg_pred, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")),
         classes_actual = cut(ddg_exp, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")))
binned_fold$ddg_pred_cat <- cut(binned_fold$ddg_pred, breaks = c(-Inf, -1.5, -.5, .5, 1.5, 2.5, 4, 5.5, 10, Inf))
binned_fold$bincode <- NA # to be filled with below loop
binned_fold$class_match <- NA
binned_fold$dataset <- "fold"

binned_bind <- loo_bind %>% 
  select(systems, mut, ddg_exp, total, error, upr, test) %>%
  rename(ddg_pred = total) %>%
  mutate(classes_pred = cut(ddg_pred, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")),
         classes_actual = cut(ddg_exp, c(-Inf, -.5, .5, Inf),
                           labels = c("stabilizing", "neutral", "destabilizing")))
binned_bind$ddg_pred_cat <- cut(binned_bind$ddg_pred, breaks = c(-1.5, -.5, .5, 1.5, 2.5, 4, 5.5, 10, Inf))

binned_bind$bincode <- NA # to be filled with below loop
binned_bind$class_match <- NA
binned_bind$dataset <- "bind"
```

3-digit bincode into a function
```{r}
create_bincode <- function(ddg, upr){
  L <- ddg - upr
  U <- ddg + upr
  if (L < -0.5 & U > 0.5){
      bincode <- "111" #any
    } else if (L > -0.5 & U < 0.5){
      bincode <- "010" #neutral
    } else if (U < -0.5){
      bincode <- "100" #stabilizing
    } else if (L > 0.5){
      bincode <- "001" #destabilizing
    } else if (U < 0.5){
      bincode <- "110" #neutral or stabilizing
    } else if (L > -0.5){
      bincode <- "011" #neutral or destabilizing
    }
  return(bincode)
}

for (i in 1:nrow(binned_fold)){ 
  binned_fold$bincode[i] <- create_bincode(binned_fold$ddg_pred[i], binned_fold$upr[i])
  if (binned_fold$classes_pred[i] == binned_fold$classes_actual[i]){
    binned_fold$class_match[i] <- "match"
  } else (binned_fold$class_match[i] <- "mismatch")
}

for (i in 1:nrow(binned_bind)){
  binned_bind$bincode[i] <- create_bincode(binned_bind$ddg_pred[i], binned_bind$upr[i])
  if (binned_bind$classes_pred[i] == binned_bind$classes_actual[i]){
    binned_bind$class_match[i] <- "match"
  } else (binned_bind$class_match[i] <- "mismatch")
}

binned_all <- bind_rows(binned_fold, binned_bind)

# Create bins on ddg_exp
binned_all$ddg_exp_cat <- cut(binned_all$ddg_exp, breaks = c(-Inf, -1.5, -.5, .5, 1.5, 2.5, 4, 5.5, Inf))
# redo binning the ddg_pred
binned_all$ddg_pred_cat <- cut(binned_all$ddg_pred, breaks = c(-Inf, -1.5, -.5, .5, 1.5, 2.5, 4, 5.5, Inf))
```


Histogram to show actual ddG distribution of our dataset:
```{r}
actual_hist <-
binned_all %>% ggplot(aes(x = ddg_exp)) +
  geom_histogram(aes(fill = classes_actual), color = "black",
                 breaks = c(-3,-1.5, -0.5, 0.5, 1.5, 2.5, 4, 5.5, 8, 16)) +
  scale_fill_manual(values = c("lightsteelblue2", "white", "rosybrown2"),
                    name = "Actual class",
                    labels = c("stabilizing", "neutral", "destabilizing")) +
  theme_classic() +
  theme(legend.position = c(.45, .5),
        legend.box.background = element_rect(),
        axis.title = element_text(size = 16), axis.text = element_text(size = 12),
        strip.text = element_text(size = 16)) +
  labs(x = expression(paste(Delta, Delta, "G")[exp]~(kcal/mol))) +
  facet_grid(~factor(dataset, levels = c("fold", "bind")))
actual_hist
```



Probably should re-do binned figures separately for fold and bind, to compare with simulation (which should be separate since the nature of intervals are different)

```{r}
class1 <- 
binned_all %>% ggplot(aes(x = ddg_pred)) +
  geom_histogram_pattern(position = "fill", color = "black",
                 breaks = c(-3,-1.5, -0.5, 0.5, 1.5, 2.5, 4, 5.5, 8, 16),
                 #breaks = c(-2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 8, 10, 12, 16),
                 pattern_color = "black", pattern_density = 0.01, pattern_spacing = 0.02,
                 aes(fill = classes_actual, pattern = class_match)) +
  scale_pattern_manual(values = c("none", "stripe"), ) +
  scale_fill_manual(values = c("lightsteelblue2", "white", "rosybrown2"),
                    name = "Actual class",
                    labels = c("stabilizing", "neutral", "destabilizing")) +
  theme_classic() +
  theme(axis.title = element_text(size = 16), axis.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.45, .3),
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[FoldX]~(kcal/mol)),
       y = "proportion") +
  guides(pattern = "none") +
  facet_grid(~factor(dataset, levels = c("fold", "bind")))
class1
```
```{r}
class2 <- 
binned_all %>% ggplot(aes(x = ddg_pred)) +
  geom_histogram(position = "fill", color = "black",
                 #breaks = c(-3, -2, -1, 0, 1, 2, 3, 4, 5, 6, 8, 10, 12, 16),
                 #breaks = c(-2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 8, 10, 12, 16),
                 breaks = c(-3,-1.5, -0.5, 0.5, 1.5, 2.5, 4, 5.5, 8, 16),
                 aes(fill = bincode)) +
  scale_fill_manual(values = c("red4", "rosybrown2", "lightsteelblue2", "white"),
                    name = "Possible class",
                    labels = c("destabilizing", "destabilizing or neutral", "stabilizing or neutral", "any")) +
  theme_classic() +
  theme(axis.title = element_text(size = 16), axis.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.47, .5), 
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[FoldX]~(kcal/mol)),
       y = "proportion") +
  facet_grid(~factor(dataset, levels = c("fold", "bind")))
class2
```
```{r}
ggarrange(actual_hist, class1, class2,
          labels = c("A","B","C"),
          ncol = 1)
ggsave("../figures/classifications_separated.png", width = 9, height = 11)
```



Relationship between error and upper bound, as ddg_pred increases:
```{r}
binned_all %>% ggplot(aes(x = ddg_pred, y = upr)) + geom_point(aes(color = systems)) +
  facet_grid(~factor(dataset, levels = c("fold", "bind")))
```


Resampling


```{r}
bins <- sort(unique(binned_all$ddg_exp_cat))
N <- 1000 #resample size per bin
resampled <- NULL # initialize resampled df
for (i in 1:length(bins)){
  df_ori <- binned_all %>% filter(ddg_exp_cat == bins[i])
  n_rows <- nrow(df_ori)
  sample_rows <- sample(n_rows, N, replace = TRUE)
  df_resampled <- df_ori[sample_rows,]
  resampled <- bind_rows(resampled, df_resampled)
}
N_min <- min(resampled$ddg_exp) ; N_max <- max(resampled$ddg_exp)
```

```{r}
resampled %>% ggplot(aes(x = ddg_exp)) +
  geom_histogram_pattern(position = "fill", color = "black",
                 breaks = c(N_min, -1.5, -0.5, 0.5, 1.5, 2.5, 4, 5.5, N_max),
                 pattern_color = "black", pattern_density = 0.01, pattern_spacing = 0.02,
                 aes(fill = classes_pred, pattern = class_match)) +
  scale_pattern_manual(values = c("none", "stripe"), ) +
  scale_fill_manual(values = c("lightsteelblue2", "white", "rosybrown2"),
                    name = "Predicted class",
                    labels = c("stabilizing", "neutral", "destabilizing")) +
  theme_classic() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.85, .3),
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[exp]~(kcal/mol)),
       y = "proportion") +
  guides(pattern = "none") +
  facet_grid(~factor(dataset, levels = c("fold", "bind")))

class1_resampled <- 
resampled %>% ggplot(aes(x = ddg_pred)) +
  geom_histogram_pattern(position = "fill", color = "black",
                 breaks = c(-3, -1.5, -0.5, 0.5, 1.5, 2.5, 4, 5.5, 8, 16),
                 pattern_color = "black", pattern_density = 0.01, pattern_spacing = 0.02,
                 aes(fill = classes_actual, pattern = class_match)) +
  scale_pattern_manual(values = c("none", "stripe"), ) +
  scale_fill_manual(values = c("lightsteelblue2", "white", "rosybrown2"),
                    name = "Actual class",
                    labels = c("stabilizing", "neutral", "destabilizing")) +
  theme_classic() +
  theme(axis.title = element_text(size = 16), axis.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.85, .3),
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[FoldX]~(kcal/mol)),
       y = "proportion") +
  guides(pattern = "none") +
  facet_grid(~factor(dataset, levels = c("fold", "bind")))
class1_resampled
```
```{r}
class2_resampled <- 
resampled %>% ggplot(aes(x = ddg_pred)) +
  geom_histogram(position = "fill", color = "black",
                 breaks = c(-3,-1.5, -0.5, 0.5, 1.5, 2.5, 4, 5.5, 8, 16),
                 aes(fill = bincode)) +
  scale_fill_manual(values = c("red4", "rosybrown2", "lightsteelblue2", "white"),
                    name = "Possible class",
                    labels = c("destabilizing", "destabilizing or neutral", "stabilizing or neutral", "any")) +
  # scale_fill_manual(values = c("red4", "rosybrown2", "white"),
  #                   name = "Possible class",
  #                   labels = c("destabilizing", "destabilizing or neutral", "any")) +
  theme_classic() +
  theme(axis.title = element_text(size = 16), axis.text = element_text(size = 12),
        strip.text = element_text(size = 16),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = c(.85, .3), 
        legend.box.background = element_rect()) +
  labs(x = expression(paste(Delta, Delta, "G")[FoldX]~(kcal/mol)),
       y = "proportion") +
  facet_grid(~factor(dataset, levels = c("fold", "bind")))
class2_resampled
```

```{r}
ggarrange(class1_resampled, class2_resampled,
          labels = c("A","B"),
          ncol = 1)
ggsave("../figures/classifications_resampled.png", width = 9, height = 7.5)
```

```{r}
table(binned_all$ddg_exp_cat, binned_all$ddg_pred_cat)
```

```{r}
table(resampled$ddg_exp_cat, resampled$ddg_pred_cat)
```




