---
title: "Error bias correction"
author: "Yesol Sapozhnikov"
date: '2022-07-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What this script does

FoldX-calculated ddG does not correlate with ddG_exp well. The slope of the fitted line of scatterplots can range from 0.2 to 0.6 in subsets of my dataset, and intercept around 0.4. The relationship varies between fold dataset and bind dataset. This causes a systemic bias in signed error in which negative ddG tends to be predicted higher (more destabilizing) and larger ddG tends to be predicted lower (underestimated.)

This script attempts to correct the bias and repeat the error model search process.


## Data and model search

```{r, include=FALSE}
library(tidyverse)
library(leaps)

source("functions.R")

# Full tables:
fold <- read.csv("../outputs/table_full_f.csv")
bind <- read.csv("../outputs/table_full_b.csv")
fold_tr <- read.csv("../outputs/tr_table_full_f.csv")
bind_tr <- read.csv("../outputs/tr_table_full_b.csv")

# selected models from stepwise & best subset selection
mods <- readRDS("../outputs/models.rds")

# Separate dataframes for correction
fold2 <- fold
bind2 <- bind
```


Obtain corrected ddG_pred 
```{r}
fold2$total <- correct_bias("ddg_exp", "total", fold2)
bind2$total <- correct_bias("ddg_exp", "total", bind2)

```


Modify datasets with corrected "error" column
```{r}
fold2$error <- abs(fold2$total - fold2$ddg_exp)
bind2$error <- abs(bind2$total - bind2$ddg_exp)

fold_tr$error <- fold2$error
bind_tr$error <- bind2$error
```

Check correction:
```{r}
par(mfrow = c(2,2))
plot(fold$ddg_exp, fold$total, pch = 20, main = "ddG_exp vs ddG_pred, Fold")
abline(0,1, col = "blue")
plot(fold2$ddg_exp, fold2$total, pch = 20, main = "ddG_exp vs ddG_pred_corrected, Fold")
abline(0,1, col = "blue")
plot(bind$ddg_exp, bind$total, pch = 20, main = "ddG_exp vs ddG_pred, Bind")
abline(0,1, col = "blue")
plot(bind2$ddg_exp, bind2$total, pch = 20, main = "ddG_exp vs ddG_pred_corrected, Bind")
abline(0,1, col = "blue")
```

Model search (exhaustive)
```{r}
mod_corr_f <- run_subset(fold_tr)
mod_corr_b <- run_subset(bind_tr)
```

Cross-validation
```{r, include = FALSE}
loo_fold <- loocv(formula(mods$mod_subset_f), fold)
loo_bind <- loocv(formula(mods$mod_subset_b), bind)

loo_fold2 <- loocv(formula(mod_corr_f), fold2)
loo_bind2 <- loocv(formula(mod_corr_b), bind2)

# Table of values for comparison
data.frame(correction = c("not corrected", "corrected"),
           fold_coverage = c(loo_fold$coverage, loo_fold2$coverage),
           fold_median_width = c(loo_fold$med_upr_bd, loo_fold2$med_upr_bd),
           bind_coverage = c(loo_bind$coverage, loo_bind2$coverage),
           bind_median_width = c(loo_bind$med_upr_bd, loo_bind2$med_upr_bd))

```

With RMSE
```{r}
rmse1_fold <- sqrt(mean((loo_fold$df$total-loo_fold$df$ddg_exp)^2))
rmse2_fold <- sqrt(mean((loo_fold2$df$total-loo_fold2$df$ddg_exp)^2))
rmse1_bind <- sqrt(mean((loo_bind$df$total-loo_bind$df$ddg_exp)^2))
rmse2_bind <- sqrt(mean((loo_bind2$df$total-loo_bind2$df$ddg_exp)^2))

data.frame(correction = c("not corrected", "corrected"),
           fold_coverage = c(loo_fold$coverage, loo_fold2$coverage),
           fold_median_width = c(loo_fold$med_upr_bd, loo_fold2$med_upr_bd),
           fold_rmse = c(rmse1_fold, rmse2_fold),
           bind_coverage = c(loo_bind$coverage, loo_bind2$coverage),
           bind_median_width = c(loo_bind$med_upr_bd, loo_bind2$med_upr_bd),
           bind_rmse = c(rmse1_bind, rmse2_bind))
```



```{r}
loo_fold2$formula
loo_bind2$formula
```



Plot signed error and compare with before the correction
```{r}
loo_fold$df$signed_error <- loo_fold$df$total-loo_fold$df$ddg_exp
loo_bind$df$signed_error <- loo_bind$df$total-loo_bind$df$ddg_exp
loo_fold2$df$signed_error <- loo_fold2$df$total-loo_fold2$df$ddg_exp
loo_bind2$df$signed_error <- loo_bind2$df$total-loo_bind2$df$ddg_exp

# combine fold and bind datasets for plotting convenience
loo_fold$df$energy <- "fold"
loo_bind$df$energy <- "bind"
loo_all <- bind_rows(loo_fold$df, loo_bind$df)
loo_all$energy <- factor(loo_all$energy, levels = c("fold", "bind"))

loo_fold2$df$energy <- "fold"
loo_bind2$df$energy <- "bind"
loo_all2 <- bind_rows(loo_fold2$df, loo_bind2$df)
loo_all2$energy <- factor(loo_all2$energy, levels = c("fold", "bind"))

signed_err_compare1 <- ggplot(loo_all, aes(ddg_exp, signed_error, color = test)) +
  geom_point() +
  theme_bw() +
  labs(title = "Empirical ddG and signed error by coverage") +
  xlab(expression(paste(Delta, Delta, "G")[exp])) + 
  ylab(expression(paste(Delta, Delta, "G")[pred] - paste(Delta, Delta, "G")[exp])) +
  facet_wrap(~energy) +
  scale_color_discrete(name = "Coverage", labels = c("miss", "covered"), type = c("red", "black"))
signed_err_compare2 <- ggplot(loo_all2, aes(ddg_exp, signed_error, color = test)) +
  geom_point() +
  theme_bw() +
  labs(title = "Empirical ddG and signed error (after correction) by coverage") +
  xlab(expression(paste(Delta, Delta, "G")[exp])) + 
  ylab(expression(paste(Delta, Delta, "G")[pred_corr] - paste(Delta, Delta, "G")[exp])) +
  facet_wrap(~energy) +
  scale_color_discrete(name = "Coverage", labels = c("miss", "covered"), type = c("red", "black"))
```


```{r}
signed_err_comparison <- gridExtra::grid.arrange(signed_err_compare1, signed_err_compare2, ncol = 1)
ggsave(signed_err_comparison, file = "../figures/signed_err_compare_correction.png", width = 7, height = 7)
gridExtra::grid.arrange(signed_err_compare1, signed_err_compare2, ncol = 1)
```

Plot ddG_exp and ddg_pred (with/without correction) with upr bd - does width vary in certain regions?
Also how does correction spread the deviations from 1:1 line?

Combine original and corrected dataframes for better visual comparison
```{r}
loo_all$correction <- "original"
loo_all2$correction <- "corrected"
loo_all12 <- bind_rows(loo_all, loo_all2)
loo_all12$correction <- factor(loo_all12$correction, levels = c("original", "corrected"))
ggplot(loo_all12, aes(ddg_exp, total, color = upr)) +
  geom_point() + geom_abline(slope = 1, intercept = 0, color = "blue") +
  scale_color_viridis_c(limits = c(0, 10)) + 
  theme_bw() +
  labs(title = "Empirical ddG vs predicted ddG \nComparison of uncorrected and corrected ddG_pred") +
  xlab(expression(paste(Delta, Delta, "G")[exp])) + 
  ylab(expression(paste(Delta, Delta, "G")[pred])) +
  facet_grid(rows = vars(energy), cols = vars(correction))
#ggsave("../figures/empricial_ddg_vs_pred_ddg_correction_comparison.png", width = 7, height = 7)
```

Histogram of deviations from 1:1 line
```{r}
ggplot(loo_all12, aes(x = error)) +
  geom_histogram(bins = 50) + 
  coord_cartesian(xlim = c(0, 7)) +
  theme_bw() +
  facet_grid(rows = vars(energy), cols = vars(correction)) +
  xlab("magnitude of error (deviations from 1:1 line")
ggsave("../figures/histogram_error_correction_comparison.png", width = 7, height = 5)
```







Try correction after excluding abberant systems
```{r}
# Another set of dataframes to test exclusion of outlier systems
fold3 <- fold[fold$systems != "1vqb",]
bind3 <- bind[bind$systems != "1ppf",]

fold3$total <- correct_bias("ddg_exp", "total", fold3)
bind3$total <- correct_bias("ddg_exp", "total", bind3)

fold3$error <- abs(fold3$total - fold3$ddg_exp)
bind3$error <- abs(bind3$total - bind3$ddg_exp)

fold3_tr <- fold3 %>% select(-c("ddg_exp","systems","mut","AAfrom","AAto","chain","residue","ch_res","ACC"))
bind3_tr <- bind3 %>% select(-c("ddg_exp","systems","mut","AAfrom","AAto","chain","residue","ch_res","ACC"))

fold3_tr$error <- fold3$error
bind3_tr$error <- bind3$error

mod_corr_f3 <- run_subset(fold3_tr)
mod_corr_b3 <- run_subset(bind3_tr)

loo_fold3 <- loocv(formula(mod_corr_f3), fold3)
loo_bind3 <- loocv(formula(mod_corr_b3), bind3)

rmse3_fold <- sqrt(mean((loo_fold3$df$total-loo_fold3$df$ddg_exp)^2))
rmse3_bind <- sqrt(mean((loo_bind3$df$total-loo_bind3$df$ddg_exp)^2))

data.frame(correction = c("not corrected", "corrected", "excl-corrected"),
           fold_coverage = c(loo_fold$coverage, loo_fold2$coverage, loo_fold3$coverage),
           fold_median_width = c(loo_fold$med_upr_bd, loo_fold2$med_upr_bd, loo_fold3$med_upr_bd),
           fold_rmse = c(rmse1_fold, rmse2_fold, rmse3_fold),
           bind_coverage = c(loo_bind$coverage, loo_bind2$coverage, loo_bind3$coverage),
           bind_median_width = c(loo_bind$med_upr_bd, loo_bind2$med_upr_bd, loo_bind3$med_upr_bd),
           bind_rmse = c(rmse1_bind, rmse2_bind, rmse3_bind))
```




Transformation matrix:
Try rotation and vertical shear

```{r}
theta <- -pi/4 #angle of rotation
A_rot <- matrix(c(cos(theta), -sin(theta), sin(theta), cos(theta)), 2,2)
s <- 0.5 #scale
A_shear <- matrix(c(1,s,0,1), 2,2)
ori_coor <- matrix(c(fold$ddg_exp, fold$total), byrow = TRUE, nrow = 2)
transformed_coor_rot <- A_rot %*% ori_coor
transformed_coor_shear <- A_shear %*% ori_coor
```

Get separate dataframes
```{r}
fold4 <- fold
fold5 <- fold
fold4$total <- transformed_coor_rot[2,]
fold5$total <- transformed_coor_shear[2,]
```

Plot
```{r}
test_fold1 <- fold %>% select(systems, ddg_exp, total) %>% mutate(transf = "original")
test_fold4 <- fold4 %>% select(systems, ddg_exp, total) %>% mutate(transf = "rotation")
test_fold5 <- fold5 %>% select(systems, ddg_exp, total) %>% mutate(transf = "shear")
test_fold <- bind_rows(test_fold1, test_fold4, test_fold5)
ggplot(test_fold, aes(ddg_exp, total)) +
  geom_point() + geom_abline(slope = 1, intercept = 0, color = "blue") +
  theme_bw() +
  labs(title = "Empirical ddG vs predicted ddG \nComparison of transformed ddG_pred") +
  xlab(expression(paste(Delta, Delta, "G")[exp])) + 
  ylab(expression(paste(Delta, Delta, "G")[pred])) +
  facet_wrap(~transf)
```










